Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-14 paper/agree.tex
Added Meaning subsection and posed a few questions therein

At this point it is important to state the meaning of a high-assurance component more formally. Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component, such as what might be sent over a wire by the communication fabric. An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL. As such, it binds \emph{L-values} to strings where an L-value (a concept from compilers) represents anything that can appear on the left-hand side of an assignment. In the context of this paper, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc. The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.

Let $s$ be the AGREE contract specification for a high-assurance component. The function, $\konst{eval}\ s\ \theta\dots$, evaluates $s$ on the given stream of total environments, $\theta\dots$, and returns \konst{true} if the specification is invariant along the entire stream and \konst{false} otherwise. The meaning of invariance for the specification merits some discussion. A guarantee has its intuitive meaning: the truth value of the associated expression at every step of the stream must be preserved. An eq-statement, however, is treated as a guarantee rather than a computation that adds a new binding to the environment. It asserts an equivalence between the named value and its associated expression. That equivalence must be preserved at every step of the stream.

The set $\Theta_s$ is the universe of environment bindings for the given specification $s$. 
The meaning of an AGREE contract specification is now defined as follows.
%
\[
  \mathcal{L}(s) = \{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots\}  
\]
%
The meaning is the prefix closed set of finite environment streams on which the specification is invariant. Each environment in the stream is total, meaning that it binds all L-values in the scope if the specification.

>>>>

\subsection{Meaning of AGREE specifications}

At this point it is important to state the meaning of a high-assurance
component more formally. Assume that the inputs to an AGREE contract
specification are raw strings (e.g., contiguous sequences of bytes)
for each named port of the component, such as what might be sent over
a wire by the communication fabric. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, divides the raw string into
its constituent parts as given by its data definition in AADL. As
such, it binds \emph{L-values} to strings where an L-value (a concept
from compilers) represents anything that can appear on the left-hand
side of an assignment. In the context of this paper, an $\lval$ is any
reference to a named port, field in a record, entry in an array, local
value defined by an eq-statement etc. The environment $\theta$ maps
any L-value in the scope of the AGREE contract specification to its
associated raw representation as a subset of bytes in a contiguous
string.\footnote{Might want to discuss why we want to stay down at the
representation level.}

[\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
generate a contig type for the inputs plus state vars.} ]

[Question: what is a \emph{total} environment? One that supplies at
least the bindings in $\theta$ needed to evaluate the spec?]

Let $s$ be the AGREE contract specification for a high-assurance
component. The function $\konst{eval}\; s\; \theta\dots$ evaluates $s$
on the given stream of total environments, $\theta\dots$, and
returns \konst{true} if the specification is invariant along the
entire stream and \konst{false} otherwise. The meaning of invariance
for the specification merits some discussion. A guarantee has its
intuitive meaning: the truth value of the associated expression at
every step of the stream must be preserved. An eq-statement, however,
is treated as a guarantee rather than a computation that adds a new
binding to the environment. It asserts an equivalence between the
named value and its associated expression. That equivalence must be
preserved at every step of the stream.

The set $\Theta_s$ is the universe of environment bindings for the
given specification $s$.  The meaning of an AGREE contract
specification is now defined as follows.
%
\[
  \mathcal{L}(s) = \set{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots}
\]
%
The meaning is the prefix closed set of finite environment streams on
which the specification is invariant. Each environment in the stream
is total, meaning that it binds all L-values in the scope if the
specification.
